---
title: "Stat184 Final Project-Claire Jablonski "
output: html_notebook
---
```{r include=FALSE} 
library(DataComputing)
library(tidyr)
library(mosaicData)
library(rvest)
library(dbplyr)
library(dplyr)
library(tidyverse)
library(party)

```
My final projects involves investigating a set of Thanksgiving 2015 data. Participants were asked 23 questions about their Thanksgiving celebrations & traditions, Thanksgiving festivities including Black Friday plans, and describing their background. From exploring and manipulating the data, I was able to learn more about all respondents and draw conconclusions about groups of people. The purpose of my project was to investigate relationships between categorical data. It is important to note that for some of investigations I truncated the range to only include the lower bound and not the upper bound.
```{r}
# Reading Thanksgiving Data with read.csv()
myURL <- "/Users/clairejablonski/Downloads/data-master/thanksgiving-2015/thanksgiving-2015-poll-data.csv"
ThanksgivingData <- read.csv(myURL) 
```

I began by renaming variables and removing variables that were not needed. I created a new data table called TidyThanksgiving which presented the data with nicer variable names and cleaner set up. 
```{r}
TidyThanksgivingData <-
  ThanksgivingData %>%
  rename(Respondent = RespondentID, CelebrateThanksgiving = Do.you.celebrate.Thanksgiving., MainDish = What.is.typically.the.main.dish.at.your.Thanksgiving.dinner., MainDishCooked = How.is.the.main.dish.typically.cooked., TypeofStuffing = What.kind.of.stuffing.dressing.do.you.typically.have., TypeofCranberrySauce = What.type.of.cranberry.saucedo.you.typically.have., Gravy = Do.you.typically.have.gravy., PrayatMeal = Do.you.typically.pray.before.or.after.the.Thanksgiving.meal., TravelDistance = How.far.will.you.travel.for.Thanksgiving., WatchMacysParade = Will.you.watch.any.of.the.following.programs.on.Thanksgiving..Please.select.all.that.apply....Macy.s.Parade, KidsTableAgeCutoff = What.s.the.age.cutoff.at.your..kids..table..at.Thanksgiving., Sex = What.is.your.gender., HouseholdCombinedSalary = How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year., LivingArea = How.would.you.describe.where.you.live. ) %>%
  select(-How.is.the.main.dish.typically.cooked....Other..please.specify., -What.kind.of.stuffing.dressing.do.you.typically.have....Other..please.specify., -What.type.of.cranberry.saucedo.you.typically.have....Other..please.specify., -Which.of.these.side.dishes.aretypically.served.at.your.Thanksgiving.dinner..Please.select.all.that.apply....Other..please.specify..1, -Which.of.these.side.dishes.aretypically.served.at.your.Thanksgiving.dinner..Please.select.all.that.apply....Other..please.specify., -Which.type.of.pie.is.typically.served.at.your.Thanksgiving.dinner..Please.select.all.that.apply....Other..please.specify., -Which.type.of.pie.is.typically.served.at.your.Thanksgiving.dinner..Please.select.all.that.apply....Other..please.specify..1, -Which.of.these.desserts.do.you.typically.have.at.Thanksgiving.dinner..Please.select.all.that.apply......Other..please.specify., -Which.of.these.desserts.do.you.typically.have.at.Thanksgiving.dinner..Please.select.all.that.apply......Other..please.specify..1)
```


```{r}
# head() was used to show a snapshot of the data table
head(TidyThanksgivingData) 
```

# Who celebrates Thanksgiving the most?
After creating a tidy data table, I wanted to know exactly how many of the responents celebrate Thanksgiving. To do this, I filtered out those that do celebrate Thanksgiving and found that 980 out of the 1058 people celebrate Thanksgiving. 
```{r}
TidyThanksgivingData %>%
  filter(CelebrateThanksgiving == "Yes") %>%
  summarise(total = n())
```

Next, I was interested in how many females and males responded to this survey.Grouping by sex shows that there were 544 women who completed the survey and 481 males that completed the survey. 33 people choose not to record their sex and have been removed from the information provided below. 
```{r}
TidyThanksgivingData %>%
  group_by(Sex) %>%
  filter(Sex != "") %>% # Removes anyone who did not list their sex 
  summarise(total = n())

```

I then created a data table called "Celebrate" that showed the split betwen sex and age group for those that do celebrate Thanksgiving. What I was interested in learning was if there was a big difference between sexes for each age group.By creating a dot plot it was easy to see the difference between the two sexes. There were more females than males for every age category.It is not suprising that there are more females that celebrate Thanksgiving in each category because more females completed the survey than males. 

It was also interesting to see that females 60+ celebrate Thanksgiving the most with a total of 144 people celebrating. Males aged 45-59 celebrate Thanksgiving the most with 127 people celebrating. The biggest difference between the number of females and males that celebrate Thanksgiving is for 60+ people. The smallest difference between the number of females and males that celebrate Thanksgiving is for the people between 45-59 age range.
```{r}
Celebrate <-
  TidyThanksgivingData %>%
  filter(CelebrateThanksgiving =="Yes") %>%
  filter(Sex != "")%>% 
  group_by(Sex,Age) %>%
  summarise(total = n()) # summarise() tells the total number of people that complete each aspect of a category
  
Celebrate %>%
  ggplot(aes(x= Age, y=total)) +
  geom_point(aes(color = Sex))
```

# What is the relationship between Household Combined Salaries and whether you work in Retail?
Next, I wanted to investigate the relationship between household combined salary ranges with if the person worked in retail or not. I created a new data table called "RetailSalary" to compare people who work in retail  with those who don't work in retail based on what their household combined salaries were. The graph below shows that more people did not work in retail than did work in retail. The graph also shows that people who did not work in retail had higher household combined salaries. 

It was interesting to see that the biggest difference in amount of people that worked in retail and did not work in retail was for the household combined total salary of $25,000 to $49,000. For both working in retail and not working in retail the most amount of people fall in the $25,000 to $49,000 range for household combined salary.It is also intersting to point out that as the household combined salary ranges increase the difference in the total amount of people that work in retail and do not work in retail decreases. This means that the range between the two groups gets smaller as the household combined salary increases.
```{r}
# Created new data table called RetailSalary where I grouped by if they worked in retail or not and their household combined salary
RetailSalary <-
  TidyThanksgivingData %>%
  group_by(Do.you.work.in.retail.,HouseholdCombinedSalary ) %>%
  filter(Do.you.work.in.retail. != "") %>%                                                       # Filter out those who did not respond to the question Do you work in retail
  filter(HouseholdCombinedSalary != "", HouseholdCombinedSalary != "Prefer not to answer") %>%   # Filter out those that did not share their combined household salary or perfered not to answer
  summarise(total = n()) %>%
  arrange(desc(total))


RetailSalary %>%
  ggplot(aes(x = HouseholdCombinedSalary, y = total)) +
  geom_point(aes( color = Do.you.work.in.retail.)) +
  theme(axis.text.x = element_text(angle = 30, hjust =1))
# Used theme() to have the x variable written at a slant so it can be visible on the graph
```

# What is the most popular Main Dish cooked in all US Regions? How is it cooked?
From looking at the original TidyThanksgivingData, I was interested in comparing different regions people live in and their main dish for Thanksgiving and how it is cooked. I expected that turkey would be the main dish for all regions and that baking the turkey would be the most popular option.I created a data table "USRegionMainDish" that explored these variables.From the table we can see that the most amount of people live in the South Atlantic and bake turkey.From looking at the chart it is difficult to see if baking or roasted is most popular cooking method and if turkey is the most popular food choice. To try and clear this up a plot was created that created a facet for each US Region exploring the differences in the food cooked and the cooking methods used.
```{r}
USRegionMainDish <-
  TidyThanksgivingData %>%
  group_by(US.Region, MainDish, MainDishCooked) %>%
  filter(US.Region != "") %>%
  filter(MainDish !="Other (please specify)") %>%
  filter(MainDish != "") %>%
  filter(MainDish != "I don't know") %>%
  filter(MainDishCooked !="Other (please specify)") %>%
  filter(MainDishCooked != "I don't know") %>%
  summarise(total = n())

USRegionMainDish %>%
  arrange( desc(total)) # Arranging the data in descending order according to total amount
```
From this graph it is now obvious that in each region that turkey is more served than anyother dish. What we cannot really tell from this graph is the way that most people cook their main dish it really varies from region to region and there is no distinct pattern seen from it. Other factors may have an influenece like age, sex, or income. Let us explore what age has to do with the style in how people cook their turkey.
```{r}
USRegionMainDish %>%
  ggplot(aes(x = MainDish , y = total )) +
  geom_point( aes(shape = MainDishCooked)) +
  facet_wrap( ~ US.Region) +
  theme(axis.text.x = element_text(angle = 30, hjust =1))

```
From this graph we get a better visaulization with color seeing the split in the amount of people in each region that either baked, fried, or roasted their turkey. We can see that all 9 US Regions bake their turkey while only 6 US Regions fry their turkey and 8 regions roast their turkey. It is obvious to tell from this graph that frying turkey is the least popular choice among all regions. West North Centeral is the only US Region that does not roast their turkey. It is very close for all other US regions between the amount of people that bake their turkey or roast it. It is interesting to note that there is larger spread between the regions in who bakes their turky compared to the more compact spread of those that roast their turkey. 
```{r}
USRegionMainDish %>%
  filter(MainDish == "Turkey" ) %>%
  ggplot(aes(x = MainDishCooked , y = total )) +
  geom_point(aes(color = US.Region))
  
```
This shows that Baked and Roasted are the most popular cooking type for the main dish. It is important to notice that the older range of people are the most popular in these types more than younger people because from what we saw before in the data it makes sense because there are more older people that complete this survey that celebrate thanksgiving compred to younger generations of people.
```{r}
MainDishCookedAge <-
  TidyThanksgivingData %>%
  group_by(Age, MainDishCooked) %>%
  summarise(total = n()) %>%
  filter(MainDishCooked !="", MainDishCooked != "Other (please specify)", MainDishCooked != "I don't know", Age != "") %>%
  arrange( desc(total) )
head(MainDishCookedAge)
```


# What is the mean Age and Salary of Surveyors?
From looking at the data, I was really interested in seeing the average age and salary people earned in the TidyThanksgivingData. To do this I created a new HouseholdSalary data table where I truncated the age and and household combined salary. To truncate the household combined salary, I created a new variable called salary that only contained the first value of the range of values. It was interesting exploring these variables and seeing that the average truncated salary was $74,077.62. The average truncated age that responded to this survey was 42 years old. It is important to see that the average truncated age was 42 showing that mostly older people responded to this survey.
```{r}
HouseholdSalary <-
  TidyThanksgivingData %>%
  mutate( HouseholdCombinedSalary=gsub("[$,%]|to|and up","",HouseholdCombinedSalary)) %>%
  # Removed $ and the words that indicated range for HouseholdCombinedSalary
  mutate( Age=gsub("[-, 29, 44, +]","", Age)) %>%
  # Removed the range values for Age variable 
  mutate(Age = as.numeric(Age)) %>%
  # Made sure Age was a numeric variable
  mutate(Salary=if_else(condition = (HouseholdCombinedSalary == "Prefer not answer"), true = as.character(NA), false = str_extract(string = HouseholdCombinedSalary, pattern = "^[[:digit:]]+" ))) %>%
  # Create a new variable Salary that only contains the first value of the range of values for the HouseholdCombinedSalary
  mutate(Salary = as.numeric(Salary))
```

```{r}
HouseholdSalary %>%
summarise(mean(Salary, na.rm = TRUE))
```
```{r}
HouseholdSalary %>%
  summarise(mean(Age, na.rm = TRUE))
```
With this truncated data, I was also interested in creating a tree graph to display Age and Salary. I think it is interesting and it makes sense that as you get older your salary will increase. One thing to take note of is that no one took the survey that was under 18 years old so it does not make sense that when you are under 18 your early salary will be $42,660.82. I also think it is interesting to see that people between the ages of 18 and 30 earn a yearly salary of $64,466.67 which is around $10,000 less than the mean salary for all people surveyed. Tree graphs make it easy to make connections between data!
```{r}
TidyHouseholdSalary <-
  HouseholdSalary %>%
  filter(Age !="NA") %>%
  filter(Salary !="NA") %>%
  filter(Sex != "")
MyModel <-
  party::ctree(Salary ~ Age + Sex, data = TidyHouseholdSalary)
plot(MyModel, type = "simple")
```


# Are Sex and Praying Before a Meal correlated?
Working with all categorical variables can be tough when trying to analyze the data. But looking at my data, I was interested in seeing if there was a correlation between praying before Thanksgiving dinner and sex. First, I created a new data table called "PrayBeforeMeal". This data table summarized the total amount of males and females who do and do not pray before Thanksgiving dinner. It was interesting to see that there were very similar totals for males and females in who prayed and who did not pray before Thanksgiving.
```{r}
PrayBeforeMeal <-
TidyThanksgivingData %>%
  group_by(PrayatMeal, Sex ) %>%
  filter(PrayatMeal != "") %>%
  filter(Sex != "") %>%
  mutate( Age=gsub("[-, 29, 44, +]","", Age)) %>%
  mutate(Age = as.numeric(Age)) %>%
  summarise(total = n())

PrayBeforeMeal %>%
  arrange( desc(total))
```

```{r}
PrayBeforeMeal %>%
  ggplot(aes( x = PrayatMeal, y= total)) +
  geom_point(aes(shape = Sex))
```


I then created a matrix where I correctly labeld gender and if they prayed or not. With creating a matrix I was able to see the proportion for each category and how much that proportion made up the entire total. This format allowed me to do a Chi Square Test where I discovered the Chi Square value is 0.20362 and the p-value is 0.6518. Because the p-value is greater than 0.05 we reject the null hypothesis which means there is no relationship between gender and praying before Thanksgiving dinner. This makes sense from looking at the graph above as the difference between males and females is not big. 
```{r}
PrayBeforeMeals <- matrix(c(174,152,341,280) ,nrow=2,byrow=TRUE)
# Created a matrix with the total values for each category
PrayBeforeMeals
```

```{r}
# Put names on the matrix so it would be easier to read
dimnames(PrayBeforeMeals) <- list(c("No","Yes"),c("Female","Male"))
names(dimnames(PrayBeforeMeals)) <- c("Pray","Gender")
PrayBeforeMeal
```

```{r}
# Found the total amount of males and females that responded to this question
tot <- sum(PrayBeforeMeals)
tot
# Found the proportion for each category 
PrayBeforeMeals/tot
```

```{r}
# Did a Chi Square test to see if gender and praying before a meal are correlated
prop.test(PrayBeforeMeals, correct = F)
```


# Does Age and Living Area have an impact on Kids Table Cuttoff?
Lastly, I was interested in seeing how the age of the respondent, the area they lived in, and their families kid table age cut off differ. I know in my household it is always the grandchildren who are forced to the kids table ( no matter how old we are )! To do this I decided to create a jitter plot. A jitter plot adds a small amount of randcom variation to the location of each point, and is useful plot for handling overplotting. From the jitter plot, I was able to see that for all age ranges in a suburban living area is it most popular to have the kids table cut off be at 10 years old. People aged 60 from the suburban area have the highest total amount of people that have their kids cut off at 10 years old. It is interesting to see the variety between all different living areas. It is very interesting to see that the kids table cutoff age was mostly either 10 or 21, values in between occured but were not as popular for all three living areas.

It is important to note that the data was truncated to only include the first value for the age range so a jitter plot could be created!
```{r}
KidsTableCutoff <-
  TidyThanksgivingData %>%
  mutate( KidsTableAgeCutoff=gsub("or|younger|older","", KidsTableAgeCutoff)) %>%
  # Removed the words or, younger, and older from the KidsTableAgeCutoff variable
  mutate( Age=gsub("[-, 29, 44, +]","", Age))
  # Removed the range values to truncate the data to only have the first value of the range

```

```{r}
KidsTableCutoff %>%
  filter(KidsTableAgeCutoff != "") %>%
  filter(Age != "") %>%
  filter(LivingArea != "") %>%
  ggplot(aes(x= Age, y = KidsTableAgeCutoff)) +
  geom_jitter(width = 0.3, height = 0.3) +
  facet_wrap( ~LivingArea)
```

# Conclusion
The purpose of this project was to explore Thanksgiving 2015 data and notice relationships between categorical variables. In each part of this project I made connections between different variables. It was hard at times to make different connections with just categorical variables so in some instances I truncated the data to only include the lower bound of a categorical variable making it a quantitative variable. Overall I enjoyed this project and learned a lot about Thanksgiving celebrations and the people that celebrate it!